On $K_2$-analogue of Serre problem for $\SL_4$

### 2018-07-31 paper.tex
Fixed statement of Lemma 2.1

For $n \geq 1$ consider the following collections of generators:
\[\XX{n} = \{ x_\alpha(\xi) \mid \xi \in B_k,\text{ for } k\leq n\}. \]
It is obvious that $\XX{n} \subseteq \XX{n+1}$, denote by $\XX{\infty}$ the union of all $\XX{n}$'s.

Let $\Phi$ be a simply-laced irreducible root system of rank $\geq 3$.
Denote by $G_m$ the group presented by generators $\mathcal{X}_m$ and the set of relations $\RR{m}$ consisting of the following
 three families of relations:
\begin{align}
 \label{eq:am} \tag{$a_m$} x_\alpha(\xi) x_\alpha(\eta) & = x_\alpha(\xi+\eta),&  \xi,\eta\in B_k,\ k\leq m;&\\
 \label{eq:bm} \tag{$b_m$} [x_\alpha(\xi), x_{\alpha'}(\eta))] &  = 1, & \text{in the case $\alpha+\alpha'\not\in\Phi\cup\{0\}$,}\\
 \nonumber                                                     &       & \xi \in B_k,\ \eta \in B_l,\ l,k\leq m;\\
 \label{eq:cm} \tag{$c_m$} [x_\alpha(\xi), x_{\alpha'}(\eta)] & = x_{\alpha+\alpha'}(N_{\alpha,\alpha'}\xi\eta), & \text{in the case $\alpha+\alpha'\in \Phi$,}\\
 \nonumber                                                    &                                                  & \xi \in B_k,\ \eta \in B_l,\ l, k, l+k \leq m.
\end{align}

The aim of this subsection is to the following analogue of~\cite[Lemma~3.3]{Tu83}. 
\begin{lemma}\label{lem:tul3.3}
 For all simply-laced root systems of rank $\geq 3$ the natural inclusion of generators induces an isomorphism $G_1 \to \St(\Phi, B).$
\end{lemma}
Unfortunately, the lemma cannot be proved by a slight modification of the original short argument of Tulenbayev.
 %for even in the linear case it works only for $\mathrm{rk}(\Phi)\geq 4$, and, in particular, is unsuitable in the case $\Phi=\mathsf{D}_\ell$.
Instead, we have to employ a variant of much longer argument of U.~Rehmann and C.~Soul{\'e}, see~\cite{Re75,RS76}.

%Denote by $F(S)$ the free group on the generating set $S$.
%It is clear that $\RR{n}\subseteq\RR{n+1}$, moreover $\langle \RR{n} \rangle = \langle \RR{\infty}\rangle \cap F(\XX{n})$.
%well, not so clear, actually :)
To prove~\cref{lem:tul3.3} it suffices to show that the natural map $\theta_m \colon G_m \to G_{m+1}$ is an isomorphism for $m\geq 1$.
Indeed, this would give the required isomorphism $G_1 \cong \St(\Phi, B)$ since $\langle \XX{\infty} \mid \RR{\infty} \rangle$ is evidently isomorphic to $\St(\Phi, B)$.

\begin{rem}
There are two reasons why we can't simply refer to~\cite{Re75} or~\cite{RS76} for the proof of~\cref{lem:tul3.3}.
The first is that Rehmann and Soul{\'e} prove that $G_m \to G_{m+1}$ is an isomorphism for $m\geq 2$, while we want to establish this fact starting from $m = 1$.
The second is that they prove their results only in the special case $B = A[t]$, moreover, they make assumptions about the ground ring $A$
(it is assumed that $A=k$ is a field in~\cite{Re75} and that $A=\mathbb{Z}$ in~\cite{RS76}).
% (this can be replaced by the weaker assumption that $A$ is additively generated by its units).
\end{rem}

>>>>

For $n \geq 1$ consider the following collections of homogeneous generators:
\[\XX{n} = \{ x_\alpha(\xi) \mid \xi \in B_k,\text{ for } k\leq n\}. \]
It is obvious that $\XX{n} \subseteq \XX{n+1}$, denote by $\XX{\infty}$ the union of all $\XX{n}$'s.

Let $\Phi$ be a simply-laced irreducible root system of rank $\geq 3$.
Fir $m\geq 2$ denote by $G_m$ the group presented by generators $\mathcal{X}_m$ and the set of relations $\RR{m}$ consisting of the following
 three families of relations:
\begin{align}
 \label{eq:am} \tag{$a_m$} x_\alpha(\xi) x_\alpha(\eta) & = x_\alpha(\xi+\eta),&  \xi,\eta\in B_k,\ k\leq m;&\\
 \label{eq:bm} \tag{$b_m$} [x_\alpha(\xi), x_{\alpha'}(\eta))] &  = 1, & \text{in the case $\alpha+\alpha'\not\in\Phi\cup\{0\}$,}\\
 \nonumber                                                     &       & \xi \in B_k,\ \eta \in B_l,\ l,k\leq m;\\
 \label{eq:cm} \tag{$c_m$} [x_\alpha(\xi), x_{\alpha'}(\eta)] & = x_{\alpha+\alpha'}(N_{\alpha,\alpha'}\xi\eta), & \text{in the case $\alpha+\alpha'\in \Phi$,}\\
 \nonumber                                                    &  & \xi \in B_k,\ \eta \in B_l,\ l, k, l+k \leq m.
 \end{align}

Now define the group $G_1$ presented by generators $\mathcal{X}_1$ and the set of relations $\RR{1}$ consisting of 
 the above families of relations $(a_1)$, $(b_1)$, $(c_1)$ with one extra family of relations
\begin{equation} \nonumber \tag{$d_1$}
[[x_{\alpha}(\xi t), x_{\beta}(\eta)],  x_{\gamma}(\zeta t)] = [[x_{\alpha}(\xi t), x_{\beta}(\eta t)],  x_{\gamma}(\zeta)], 
\end{equation}
where $\alpha, \beta, \gamma \in \Phi$ are such that $\alpha+\beta+\gamma \in \Phi$ and $\xi, \eta, \zeta \in A$.

The aim of this subsection is to demonstrate the following analogue of~\cite[Lemma~3.3]{Tu83}. 
\begin{lemma}\label{lem:tul3.3}
 Let $\Phi$ be a simply-laced root systems of rank $\geq 3$. 
 Then the natural inclusion of generators $\mathcal{X}_n \subseteq \mathcal{X}_{n+1}$ induces an isomorphism $G_n \to G_{n+1}$ for $n\geq 1$.
 Consequently, the group $\St(\Phi, B)$ is isomorphic to $G_1$.
\end{lemma}
Unfortunately, the lemma cannot be proved by a slight modification of the original short argument of Tulenbayev.
 %for even in the linear case it works only for $\mathrm{rk}(\Phi)\geq 4$, and, in particular, is unsuitable in the case $\Phi=\mathsf{D}_\ell$.
Instead, we have to employ a varition of a much longer argument of U.~Rehmann and C.~Soul{\'e}, cf~\cite{Re75,RS76}.

%Denote by $F(S)$ the free group on the generating set $S$.
%It is clear that $\RR{n}\subseteq\RR{n+1}$, moreover $\langle \RR{n} \rangle = \langle \RR{\infty}\rangle \cap F(\XX{n})$.
%well, not so clear, actually :)
To prove~\cref{lem:tul3.3} it suffices to show that the natural map $\theta_m \colon G_m \to G_{m+1}$ is an isomorphism for $m\geq 2$.
Indeed, this would give the required isomorphism $G_1 \cong \St(\Phi, B)$ since $\langle \XX{\infty} \mid \RR{\infty} \rangle$ is evidently isomorphic to $\St(\Phi, B)$.

%\begin{rem}
%There are two reasons why we can't simply refer to~\cite{Re75} or~\cite{RS76} for the proof of~\cref{lem:tul3.3}.
%The first is that we want to prove Rehmann and Soul{\'e} prove that $G_m \to G_{m+1}$ is an isomorphism for $m\geq 2$, while we want to %establish this fact starting from $m = 1$.
%The second is that they prove their results only in the special case $B = A[t]$, moreover, they make assumptions about the ground ring $A$
%(it is assumed that $A=k$ is a field in~\cite{Re75} and that $A=\mathbb{Z}$ in~\cite{RS76}).
% (this can be replaced by the weaker assumption that $A$ is additively generated by its units).
%\end{rem}
